---
title: "R Project"
author: "Austin Smith", "Nicholas Vota"
output: html_notebook
---

Sudoku Input for the time being -- may decide to design our own if time permits.

```{r}
library(sudoku)
puzzle <- fetchSudokuUK()
typeof(puzzle)
```
# Cracking a single row:
-needs to check container, row, and column for whether the multiplicity of 1-9 is 0 or 1
-0 = move to next environment (row/column) and 1 = non suitable variable
-should be hard to explicitly determine which variable, therefore a list of multiple potential variables (1-9) should be generated
```{r}
set.seed(42)
library(sudoku)
puzzle <- generateSudoku()
puzzle
#loop for dividing the board into 3x3 containers contained in an array
containers <- array(dim= c(3,3,9))
for(n in 1:3){
  if(n == 1){
    for(i in 1:3){
      containers[,,i] <- puzzle[1:3,((3*i)-2):(3*i)]
    }
  } else if(n == 2){
    for(i in 1:3){
      containers[,,i+3] <- puzzle[4:6,((3*i)-2):(3*i)]
    }
  } else{
    for(i in 1:3){
      containers[,,i+6] <- puzzle[7:9,((3*i)-2):(3*i)]
    }
  }
}
#loop for column and row arrays
columns <- array(dim=c(1,9,9))
for(n in 1:9){
  columns[,,n] <- puzzle[1:9, n]
}
rows <- array(dim=c(1,9,9))
for(n in 1:9){
  rows[,,n] <- puzzle[n, 1:9]
}
#checking for a single varible across a row
possibility_matrix <- array(dim=c(9,9,9))
for(i in 1:9){
  #spliting into container
  if(i <= 3){
    #first checking container for optimal efficiency
    if(sum(containers[,,1] == 1) == 0){
      
      if(puzzle[1,i] == 0){
        
        if(sum(rows[,,1] == 1) == 0){
          
          if(sum(columns[,,i] == 1) == 0){
            
            possibility_matrix[1,i,1] = 1
            
          } else if(sum(columns[,,i] == 1) == 1){
            
            next
            
          } else{
            
            stop("Error: Multiplicity of a variable has exceeded 1 A")
            
          }
          
        } else if(sum(rows[,,1] == 1) == 1){
          
          next
          
        } else {
      
          stop("Error: Multiplicity of a variable has exceeded 1 B")
      
        }
        
      } else{
    
        next
    
      }
      
    } else if(sum(containers[,,1] == 1) == 1){
      
      next
      
    } else{
      
      stop("Error: Multiplicity of a variable has exceeded 1 C")
      
    }
  } else if(i <= 6){
    
    if(sum(containers[,,2] == 1) == 0){
      
      if(puzzle[1,i] == 0){
        
        if(sum(rows[,,1] == 1) == 0){
          
          if(sum(columns[,,i] == 1) == 0){
            
            possibility_matrix[1,i,1] = 1
            
          } else if(sum(columns[,,i] == 1) == 1){
            
            next
            
          } else{
            
            stop("Error: Multiplicity of a variable has exceeded 1 D")
            
          }
          
        } else if(sum(rows[,,1] == 1) == 1){
          
          next
          
        } else {
      
          stop("Error: Multiplicity of a variable has exceeded 1 E")
      
        }
        
      } else{
    
        next
    
      }
      
    } else if(sum(containers[,,2] == 1) == 1){
      
      next
      
    } else{
      
      stop("Error: Multiplicity of a variable has exceeded 1 F")
      
    }
    
  } else{
    if(sum(containers[,,3] == 1) == 0){
      
      if(puzzle[1,i] == 0){
        
        if(sum(rows[,,1] == 1) == 0){
          
          if(sum(columns[,,i] == 1) == 0){
            
            possibility_matrix[1,i,1] = 1
            
          } else if(sum(columns[,,i] == 1) == 1){
            
            next
            
          } else{
            
            stop("Error: Multiplicity of a variable has exceeded 1 G")
            
          }
          
        } else if(sum(rows[,,1] == 1) == 1){
          
          next
          
        } else {
      
          stop("Error: Multiplicity of a variable has exceeded 1 H")
      
        }
        
      } else{
    
        next
    
      }
      
    } else if(sum(containers[,,3] == 1) == 1){
      
      next
      
    } else{
      
      stop("Error: Multiplicity of a variable has exceeded 1 L")
      
    }
  }
  
}
#This is this spots which can possibly be 1 (1 matrix for each variable), this will be the master matrix
possibility_matrix[,,1]
```

Okay now we know we can create a system of matrixes that will contain every possible number for each space, we just need to build this
function out to each row and each variable. After we have this matrix it should be easy to wrtie a recursive program that analyzes it to determine the answers.

Below is the full matrix:

```{r}
containers <- array(dim= c(3,3,9))
for(n in 1:3){
  if(n == 1){
    for(i in 1:3){
      containers[,,i] <- puzzle[1:3,((3*i)-2):(3*i)]
    }
  } else if(n == 2){
    for(i in 1:3){
      containers[,,i+3] <- puzzle[4:6,((3*i)-2):(3*i)]
    }
  } else{
    for(i in 1:3){
      containers[,,i+6] <- puzzle[7:9,((3*i)-2):(3*i)]
    }
  }
}
#loop for column and row arrays
columns <- array(dim=c(1,9,9))
for(n in 1:9){
  columns[,,n] <- puzzle[1:9, n]
}
rows <- array(dim=c(1,9,9))
for(n in 1:9){
  rows[,,n] <- puzzle[n, 1:9]
}
#checking for a single varible across a row
possibility_matrix <- array(0,dim=c(9,9,9))
for(k in 1:9){
  for(n in 1:9){
    for(i in 1:9){
    #spliting into container
      if(i <= 3 && n <= 3){
      #first checking container for optimal efficiency
        if(sum(containers[,,1] == k) == 0){
        
          if(puzzle[n,i] == 0){
          
            if(sum(rows[,,n] == k) == 0){
            
              if(sum(columns[,,i] == k) == 0){
              
                possibility_matrix[n,i,k] = k
              
              } else if(sum(columns[,,i] == k) == 1){
              
                next
              
              } else{
              
                stop("Error: Multiplicity of a variable has exceeded 1 A")
              
              }
            
            } else if(sum(rows[,,n] == k) == 1){
            
              next
            
            } else {
       
              stop("Error: Multiplicity of a variable has exceeded 1 B")
        
            }
          
          } else{
      
            next
      
          }
        
        } else if(sum(containers[,,1] == k) == 1){
        
          next
        
        } else{
        
          stop("Error: Multiplicity of a variable has exceeded 1 C")
        
        }
      } else if(i <= 6 && n <= 3){
      #first checking container for optimal efficiency
        if(sum(containers[,,2] == k) == 0){
        
          if(puzzle[n,i] == 0){
          
            if(sum(rows[,,n] == k) == 0){
            
              if(sum(columns[,,i] == k) == 0){
              
                possibility_matrix[n,i,k] = k
              
              } else if(sum(columns[,,i] == k) == 1){
              
                next
              
              } else{
              
                stop("Error: Multiplicity of a variable has exceeded 1 A")
              
              }
            
            } else if(sum(rows[,,n] == k) == 1){
            
              next
            
            } else {
        
              stop("Error: Multiplicity of a variable has exceeded 1 B")
        
            }
          
          } else{
      
            next
      
          }
        
        } else if(sum(containers[,,2] == k) == 1){
        
          next
        
        } else{
        
          stop("Error: Multiplicity of a variable has exceeded 1 C")
        
        }
      } else if(i <= 9 && n <= 3){
      #first checking container for optimal efficiency
        if(sum(containers[,,3] == k) == 0){
        
          if(puzzle[n,i] == 0){
          
            if(sum(rows[,,n] == k) == 0){
            
              if(sum(columns[,,i] == k) == 0){
              
                possibility_matrix[n,i,k] = k
              
              } else if(sum(columns[,,i] == k) == 1){
              
                next
              
              } else{
              
                stop("Error: Multiplicity of a variable has exceeded 1 A")
              
              }
            
            } else if(sum(rows[,,n] == k) == 1){
            
              next
            
            } else {
        
              stop("Error: Multiplicity of a variable has exceeded 1 B")
        
            }
          
          } else{
      
            next
      
          }
        
        } else if(sum(containers[,,3] == k) == 1){
        
          next
        
        } else{
        
          stop("Error: Multiplicity of a variable has exceeded 1 C")
        
        }
      } else if(i <= 3 && n <= 6){
      #first checking container for optimal efficiency
        if(sum(containers[,,4] == k) == 0){
        
          if(puzzle[n,i] == 0){
          
            if(sum(rows[,,n] == k) == 0){
            
              if(sum(columns[,,i] == k) == 0){
              
                possibility_matrix[n,i,k] = k
              
              } else if(sum(columns[,,i] == k) == 1){
              
                next
              
              } else{
              
                stop("Error: Multiplicity of a variable has exceeded 1 A")
              
              }
            
            } else if(sum(rows[,,n] == k) == 1){
            
              next
            
            } else {
        
              stop("Error: Multiplicity of a variable has exceeded 1 B")
        
            }
          
          } else{
      
            next
      
          }
        
        } else if(sum(containers[,,4] == k) == 1){
        
          next
        
        } else{
        
          stop("Error: Multiplicity of a variable has exceeded 1 C")
        
        }
      } else if(i <= 6 && n <= 6){
      #first checking container for optimal efficiency
        if(sum(containers[,,5] == k) == 0){
        
          if(puzzle[n,i] == 0){
          
            if(sum(rows[,,n] == k) == 0){
            
              if(sum(columns[,,i] == k) == 0){
              
                possibility_matrix[n,i,k] = k
              
              } else if(sum(columns[,,i] == k) == 1){
              
                next
              
              } else{
              
                stop("Error: Multiplicity of a variable has exceeded 1 A")
              
              }
            
            } else if(sum(rows[,,n] == k) == 1){
            
              next
            
            } else {
        
              stop("Error: Multiplicity of a variable has exceeded 1 B")
        
            }
          
          } else{
      
            next
      
          }
        
        } else if(sum(containers[,,5] == k) == 1){
        
          next
        
        } else{
        
          stop("Error: Multiplicity of a variable has exceeded 1 C")
        
        } 
      } else if(i <= 9 && n <= 6){
      #first checking container for optimal efficiency
        if(sum(containers[,,6] == k) == 0){
        
          if(puzzle[n,i] == 0){
          
            if(sum(rows[,,n] == k) == 0){
            
              if(sum(columns[,,i] == k) == 0){
              
                possibility_matrix[n,i,k] = k
              
              } else if(sum(columns[,,i] == k) == 1){
              
                next
              
              } else{
              
                stop("Error: Multiplicity of a variable has exceeded 1 A")
              
              }
            
            } else if(sum(rows[,,n] == k) == 1){
            
              next
            
            } else {
        
              stop("Error: Multiplicity of a variable has exceeded 1 B")
        
            }
          
          } else{
      
            next
      
          }
        
        } else if(sum(containers[,,6] == k) == 1){
        
          next
        
        } else{
        
          stop("Error: Multiplicity of a variable has exceeded 1 C")
        
        } 
      } else if(i <= 3 && n <= 9){
      #first checking container for optimal efficiency
        if(sum(containers[,,7] == k) == 0){
        
          if(puzzle[n,i] == 0){
          
            if(sum(rows[,,n] == k) == 0){
            
              if(sum(columns[,,i] == k) == 0){
              
                possibility_matrix[n,i,k] = k
              
              } else if(sum(columns[,,i] == k) == 1){
              
                next
              
              } else{
              
                stop("Error: Multiplicity of a variable has exceeded 1 A")
              
              }
            
            } else if(sum(rows[,,n] == k) == 1){
            
              next
            
            } else {
        
              stop("Error: Multiplicity of a variable has exceeded 1 B")
        
            }
          
          } else{
      
            next
      
          }
        
        } else if(sum(containers[,,7] == k) == 1){
        
          next
        
        } else{
        
          stop("Error: Multiplicity of a variable has exceeded 1 C")
        
        }
      } else if(i <= 6 && n <= 9){
      #first checking container for optimal efficiency
        if(sum(containers[,,8] == k) == 0){
        
          if(puzzle[n,i] == 0){
          
            if(sum(rows[,,n] == k) == 0){
            
              if(sum(columns[,,i] == k) == 0){
              
                possibility_matrix[n,i,k] = k
              
              } else if(sum(columns[,,i] == k) == 1){
              
                next
              
              } else{
              
                stop("Error: Multiplicity of a variable has exceeded 1 A")
              
              }
            
            } else if(sum(rows[,,n] == k) == 1){
            
              next
            
            } else {
        
              stop("Error: Multiplicity of a variable has exceeded 1 B")
        
            }
          
          } else{
      
            next
      
          }
        
        } else if(sum(containers[,,8] == k) == 1){
        
          next
        
        } else{
        
          stop("Error: Multiplicity of a variable has exceeded 1 C")
        
        }
      } else if(i <= 9 && n <= 9){
      #first checking container for optimal efficiency
        if(sum(containers[,,9] == k) == 0){
        
          if(puzzle[n,i] == 0){
          
            if(sum(rows[,,n] == k) == 0){
            
              if(sum(columns[,,i] == k) == 0){
              
                possibility_matrix[n,i,k] = k
              
              } else if(sum(columns[,,i] == k) == 1){
              
                next
              
              } else{
              
                stop("Error: Multiplicity of a variable has exceeded 1 A")
              
              }
            
            } else if(sum(rows[,,n] == k) == 1){
            
              next
            
            } else {
        
              stop("Error: Multiplicity of a variable has exceeded 1 B")
        
            }
          
          } else{
      
            next
      
          }
        
        } else if(sum(containers[,,9] == k) == 1){
        
          next
        
        } else{
        
          stop("Error: Multiplicity of a variable has exceeded 1 C")
        
        }
      }
    }
  }
}
possibility_matrix
```

Now that we have this output we can create a function to go ahead and choose numbers 
that are valid to place. This is based on:
-if the variable is the only possible variable to place in a spot.
-if other spaces in an environment (row, column, container) have a possibility of
placing that variable.
then we can add those variables to the original puzzle and repeat the process until
all spaces on the puzzle are filled.


```{r}
new_puzzle <- puzzle
for(k in 1:9){
  
  for(n in 1:9){
    
    for(i in 1:9){
      
      if(possibility_matrix[n,i,k] > 0){
        
        if(i <= 3 && n <= 3){
        
          if(sum(possibility_matrix[1:3,1:3,k] == k) == 1){
            
            new_puzzle[n,i] <- k 
            
          } else if(sum(possibility_matrix[1:9,i,k] == k) == 1){
            
            new_puzzle[n,i] <- k 
            
          } else if(sum(possibility_matrix[n,1:9,k] == k) == 1){
            
            new_puzzle[n,i] <- k
            
          } else{
            
            next
            
          }
          
        } else if(i <= 6 && n <= 3){
          
          if(sum(possibility_matrix[1:3,4:6,k] == k) == 1){
            
            new_puzzle[n,i] <- k 
            
          } else if(sum(possibility_matrix[1:9,i,k] == k) == 1){
            
            new_puzzle[n,i] <- k 
            
          } else if(sum(possibility_matrix[n,1:9,k] == k) == 1){
            
            new_puzzle[n,i] <- k
            
          } else{
            
            next
            
          }
          
        } else if(i <= 9 && n <= 3){
          
          if(sum(possibility_matrix[1:3,7:9,k] == k) == 1){
            
            new_puzzle[n,i] <- k 
            
          } else if(sum(possibility_matrix[1:9,i,k] == k) == 1){
            
            new_puzzle[n,i] <- k 
            
          } else if(sum(possibility_matrix[n,1:9,k] == k) == 1){
            
            new_puzzle[n,i] <- k
            
          } else{
            
            next
            
          }
          
        } else if(i <= 3 && n <= 6){
          
          if(sum(possibility_matrix[4:6,1:3,k] == k) == 1){
            
            new_puzzle[n,i] <- k 
            
          } else if(sum(possibility_matrix[1:9,i,k] == k) == 1){
            
            new_puzzle[n,i] <- k 
            
          } else if(sum(possibility_matrix[n,1:9,k] == k) == 1){
            
            new_puzzle[n,i] <- k
            
          } else{
            
            next
            
          }
          
        } else if(i <= 6 && n <= 6){
          
          if(sum(possibility_matrix[4:6,4:6,k] == k) == 1){
            
            new_puzzle[n,i] <- k 
            
          } else if(sum(possibility_matrix[1:9,i,k] == k) == 1){
            
            new_puzzle[n,i] <- k 
            
          } else if(sum(possibility_matrix[n,1:9,k] == k) == 1){
            
            new_puzzle[n,i] <- k
            
          } else{
            
            next
            
          }
        } else if(i <= 9 && n <= 6){
          
          if(sum(possibility_matrix[4:6,7:9,k] == k) == 1){
            
            new_puzzle[n,i] <- k 
            
          } else if(sum(possibility_matrix[1:9,i,k] == k) == 1){
            
            new_puzzle[n,i] <- k 
            
          } else if(sum(possibility_matrix[n,1:9,k] == k) == 1){
            
            new_puzzle[n,i] <- k
            
          } else{
            
            next
            
          }
        } else if(i <= 3 && n <= 9){
          
          if(sum(possibility_matrix[7:9,1:3,k] == k) == 1){
            
            new_puzzle[n,i] <- k 
            
          } else if(sum(possibility_matrix[1:9,i,k] == k) == 1){
            
            new_puzzle[n,i] <- k 
            
          } else if(sum(possibility_matrix[n,1:9,k] == k) == 1){
            
            new_puzzle[n,i] <- k
            
          } else{
            
            next
            
          }
        } else if(i <= 6 && n <= 9){
          
          if(sum(possibility_matrix[7:9,4:6,k] == k) == 1){
            
            new_puzzle[n,i] <- k 
            
          } else if(sum(possibility_matrix[1:9,i,k] == k) == 1){
            
            new_puzzle[n,i] <- k 
            
          } else if(sum(possibility_matrix[n,1:9,k] == k) == 1){
            
            new_puzzle[n,i] <- k
            
          } else{
            
            next
            
          }
        } else if(i <= 9 && n <= 9){
          
          if(sum(possibility_matrix[7:9,7:9,k] == k) == 1){
            
            new_puzzle[n,i] <- k 
            
          } else if(sum(possibility_matrix[1:9,i,k] == k) == 1){
            
            new_puzzle[n,i] <- k 
            
          } else if(sum(possibility_matrix[n,1:9,k] == k) == 1){
            
            new_puzzle[n,i] <- k
            
          } else{
            
            next
            
          }
        }
          
      } else {
        
        next
        
      }
    }
  }
  
}
new_puzzle

```